Spectroscopic confirmation of very-wide stellar binaries and large-separation comoving pairs from \textsl{Gaia DR1}

### 2017-10-02 paper/comoving-followup.tex
dispersions

\subsection{Comoving pair catalog}\label{sec:catalog}

Of the 323 observed star pairs, 12 have at least one member with an unreliable
(flagged) radial velocity measurement from the spectroscopic reduction pipeline.
For the remaining 311 pairs, we use the measured radial velocities to re-compute
the marginal likelihood ratio for the two hypotheses considered in previous work
(see Appendix~B, \citealt{Oh:2017}).
We compute two fully marginalized likelihoods (FMLs) for the data: (1) assuming
that the two stars are comoving with identical 3D velocity vectors (within some
allowed small dispersion), and (2) assuming that the two stars have independent
3D velocity vectors drawn from a prior velocity distribution.
We set the small allowed dispersion ($s$, in previous work) to $1~\kms$, and use
a mixture of three Gaussian velocity components for the prior, meant to
represent the thin disk, thick disk, and halo velocity distributions relative to
the sun (see \citealt{Oh:2017}).

@@

\section{Mixture model...}\label{sec:TODO}

% TODO: assumptions:
% - Same velocity dispersion for all ages. Could have an age-dependent
%   dispersion
% -
With the new radial velocity measurements, we construct a statistical model to
determine the probability that a given candidate comoving pair is truly
comoving.
We don't model the positional separation distribution and instead construct a
model purely in velocity-space.
The data are therefore proper motions in the ICRS reference frame, $\bs{\mu}$,
and parallax, $\varpi$, (from the \tgas\ catalog) and barycentric radial
velocity, $v_r$, for each star in a candidate comoving pair; we assume that the
sky position for each star $(\alpha, \delta)$ is known perfectly.
For brevity in the expressions below, we define $\bs{\eta} = (\mu_{\alpha^*},
\mu_\delta, v_r)$, where, e.g., $\bs{\eta}_1$ refers to the proper motion and
radial velocity data for star 1 in the pair.
To simplify some expressions, we additionally ignore any correlations between
the observed parallax and the proper motion components (but keep the covariance
structure between the two proper motion components).

Our model is expressed in terms of the magnitude of the true (unobserved) 3D
velocity difference for each pair, $\psi = \left| \bs{v}_1 - \bs{v}_2 \right|$,
and the (global) fraction of pairs that are comoving, $f$.
For the distribution of true 3D velocity difference magnitudes, $p(\psi \given
f)$, we use a mixture model with two components
\begin{eqnarray}
    p(\psi \given f) &=& f\,p_{\rm co}(\psi) + (1-f)\,p_{\rm field}(\psi)
\end{eqnarray}
where $p_{\rm co}$, the probability distribution function (pdf) of true comoving
pairs, and $p_{\rm field}$, the pdf of randomly paired field stars.\footnote{
This is actually the marginal likelihood under a model in which every candidate
comoving pair has an additional latent parameter: a binary flag, $q_n$, that
specifies which mixture component it belongs too
(\citealt{Hogg:2010,Foreman-Mackey:2014}). We later use this fact to compute the
probability that each pair is truly comoving, or, more precisely, belongs to the
mixture component $p_{\rm co}$).}
The form for both mixture components ($p_{\rm co}$ and $p_{\rm field}$) is taken
to be the probability of the magnitude of the 3D separation of two vectors under
the assumption that the two vectors were drawn independently from an isotropic,
3D Gaussian with dispersion $\sigma$.
That is,
\begin{eqnarray}
    f(\psi \,;\, \sigma) &=& \left[2\sqrt {\pi}\, \sigma^3 \right]^{-1} \,
        \psi^2 \, \ \exp\left[-\frac {\psi^2} {4\, \sigma^2} \right] \\
    \bs{v_1}, \bs{v_2} &\sim& \mathcal{N}(\bs{0}, \sigma^2\,\mat{I}_3) \\
    p_{\rm co}(\psi) &=& f(\psi \,;\, \sigma = 1~\kms)\\
    p_{\rm field}(\psi) &=& f(\psi \,;\, \sigma = \sigma_{\rm field})
\end{eqnarray}
where $\sigma_{\rm field}$ is defined below, $\mathcal{N}(\bs{m},
\mat{\Sigma})$ represents a normal distribution with mean vector $\bs{m}$ and
covariance matrix $\mat{\Sigma}$, and $\mat{I}_3$ is the 3 by 3 identity matrix.

>>>>

\section{Methods: Comoving pair probabilities}\label{sec:methods}

% \subsection{Comoving pair probabilities} \label{sec:pair-probs}

% TODO: assumptions:
% - Same velocity dispersion for all ages. Could have an age-dependent
%   dispersion
% -

Of the 323 observed star pairs, 12 have at least one member with an unreliable
(flagged) radial velocity measurement from the spectroscopic reduction pipeline.
For the remaining 311 pairs, we construct a statistical model to determine the
probability that a given candidate comoving pair is truly comoving.
For this inference, the data are the proper motions in the ICRS reference frame,
$\bs{\mu}$, and parallax, $\varpi$, (from the \tgas\ catalog) and barycentric
radial velocity, $v_r$, for each star in a candidate comoving pair; we assume
that the sky position for each star $(\alpha, \delta)$ is known perfectly.
For brevity in the expressions below, we define $\bs{\eta} = (\mu_{\alpha^*},
\mu_\delta, v_r)$, where, e.g., $\bs{\eta}_1$ refers to the proper motion and
radial velocity data for star 1 in the pair.
To simplify some expressions, we additionally ignore any correlations between
the observed parallax and the proper motion components (but keep the covariance
structure between the two proper motion components).

Our model is expressed in terms of the magnitude of the true (unobserved) 3D
velocity difference for each pair, $\psi = \left| \bs{v}_1 - \bs{v}_2 \right|$,
and the (global) fraction of pairs that are comoving, $f$.
For the distribution of true 3D velocity difference magnitudes, $p(\psi \given
f)$, we use a mixture model with two components
\begin{eqnarray}
    p(\psi \given f) &=& f\,p_{\rm co}(\psi) + (1-f)\,p_{\rm field}(\psi)
\end{eqnarray}
where $p_{\rm co}$, the probability distribution function (pdf) of true comoving
pairs, and $p_{\rm field}$, the pdf of randomly paired field stars.
Using a mixture model and inferring the mixture weight is a common trick to
avoid explicitly having to sample over the per-pair probabilities: this
expression is equivalent to the marginal likelihood under a model in which every
candidate comoving pair has an additional latent parameter (a binary flag,
$q_n$) that specifies which mixture component it belongs to
(\citealt{Hogg:2010,Foreman-Mackey:2014}).
We later use this fact to compute the probability that each pair is truly
comoving, or, more precisely, belongs to the mixture component $p_{\rm co}$.

The form for both mixture components ($p_{\rm co}$ and $p_{\rm field}$) is taken
to be the probability of the magnitude of the 3D separation of two vectors under
the assumption that the two vectors were drawn independently from an isotropic,
3D Gaussian with dispersion $\sigma$.
That is,
\begin{eqnarray}
    f(\psi \,;\, \sigma) &=& \left[2\sqrt {\pi}\, \sigma^3 \right]^{-1} \,
        \psi^2 \, \ \exp\left[-\frac {\psi^2} {4\, \sigma^2} \right] \\
    \bs{v_1}, \bs{v_2} &\sim& \mathcal{N}(\bs{0}, \sigma^2\,\mat{I}_3) \\
    p_{\rm co}(\psi) &=& f(\psi \,;\, \sigma = 1~\kms)\\
    p_{\rm field}(\psi) &=& f(\psi \,;\, \sigma = \sigma_{\rm field})
\end{eqnarray}
where $\sigma_{\rm field}$ is defined below, $\mathcal{N}(\bs{m},
\mat{\Sigma})$ represents a normal distribution with mean vector $\bs{m}$ and
covariance matrix $\mat{\Sigma}$, and $\mat{I}_3$ is the 3 by 3 identity matrix.

Dispersion ... Nordstrom 2004

\section{Old section on FMLs}
We compute two fully marginalized likelihoods (FMLs) for the data: (1) assuming
that the two stars are comoving with identical 3D velocity vectors (within some
allowed small dispersion), and (2) assuming that the two stars have independent
3D velocity vectors drawn from a prior velocity distribution.
We set the small allowed dispersion ($s$, in previous work) to $1~\kms$, and use
a mixture of three Gaussian velocity components for the prior, meant to
represent the thin disk, thick disk, and halo velocity distributions relative to
the sun (see \citealt{Oh:2017}).

@@
